---
title: "LM_Analysis"
output: html_document
date: "2024-12-28"
---

```{r setup, include=FALSE}
library(tidyverse)
library(car)         # multicollinearity 
library(lmtest)      # heteroskedasticity
library(MASS)        # stepwise regression
library(caret)       # model evaluation
library(glmnet)      # LASSO
library(AER)
library(dplyr)       # all_of()
```


FUNCTIONS_______________________
```{r}
reduce_multicollinearity <- function(data, formula, vif_threshold = 5) {
  current_model <- lm(formula, data = data)
  response_variable <- as.character(formula[[2]])
  removed_variables <- c()
  
  repeat {
    vif_values <- vif(current_model)
    max_vif <- max(vif_values)
    
    if (max_vif < vif_threshold) {
      break
    }
    
    variable_to_remove <- names(which.max(vif_values))
    
    cat("Removing variable:", variable_to_remove, "with VIF:", max_vif, "\n")
    
    removed_variables <- c(removed_variables, variable_to_remove)
    
    excluded_variables <- paste0("`", removed_variables, "`", collapse = " - ")
    updated_formula <- as.formula(paste(response_variable, " ~ . -", excluded_variables))
    current_model <- lm(updated_formula, data = data)
  }
  print(updated_formula)
  return(updated_formula)
}

check_model_assumptions <- function(model)
{
  # Multicollinearity
  vif_values <- vif(model)
  #print(vif_values)
  high_vif <- names(vif_values[vif_values > 5])
  #print(high_vif)
  
  # Homoskedasticity
  par(mfrow = c(2, 2))  
  plot(model)
  bp_result <- bptest(model)
  print(bp_result)
  
  # Normality of residuals
  qqnorm(resid(model))
  qqline(resid(model))
  s_result <- shapiro.test(resid(model))
  print(s_result)
  # Linearity
  #crPlots(model)
  
  # Summary
  print(summary(model))
}

cross_validate_lm <- function(formula, data, k = 10, seed = 111) {
  set.seed(seed) 
  
  data <- data[sample(nrow(data)), ]
  
  folds <- cut(seq(1, nrow(data)), breaks = k, labels = FALSE)
  
  errors <- numeric(k)
  r_squared <- numeric(k)
  
  for (i in 1:k) {
    test_indices <- which(folds == i, arr.ind = TRUE)
    test_data <- data[test_indices, ]
    train_data <- data[-test_indices, ]
    
    model <- lm(formula, data = train_data)
    
    predictions <- predict(model, newdata = test_data)
    
    mse <- mean((test_data[[as.character(formula[[2]])]] - predictions)^2)
    rss <- sum((test_data[[as.character(formula[[2]])]] - predictions)^2)
    tss <- sum((test_data[[as.character(formula[[2]])]] - mean(test_data[[as.character(formula[[2]])]]))^2)
    r2 <- 1 - (rss / tss)
    
    errors[i] <- mse
    r_squared[i] <- r2
  }
  
  list(
    Mean_MSE = mean(errors),
    Mean_R_squared = mean(r_squared),
    Errors_per_fold = errors,
    R_squared_per_fold = r_squared
  )
}
```

Choosing a set of predictors and testing basic assumptions, then shrinkage

Economical variables:
Energy and water earnings
Information and communication earnings
Education earnings
Accomodation and food earnings
Producers of capital goods
Retail earnings
Unemployment rate 
Vacancies count
Consumer.Price.Index
Overall.economy

Social variables:
Immigration arrivals
Immigration departures
German departures
Births
Marriages

Dependent variables
"Total.offences" - bp fails
"Offences.against.life" - OK
"Murder" - ok
"Criminal.offenses.against.sexual.self.determination.in.total" - bp fails
"Robberies.in.apartments" - ok
"Total.theft" - ok

```{r}
var <- "Total.offences"
# Load and clean data
data_monthly <- read.csv("monthly.csv")
data_cleaned <- na.omit(data_monthly)

#print(colnames(data_cleaned))
# Select columns + dependent column (var)
data_selected <- data_cleaned %>%
  dplyr::select(
    Total.offences,
    Unemployment.rate,
    Vacancies.unfilled,
    Births_Total_x,
    Marriages.per.1000.inhabitants_x,
    Germans_Total_Departures.to.foreign.countries,
    Foreigners_Total_Arrivals.from.foreign.countries,
    Foreigners_Total_Departures.to.foreign.countries,
    Turnover.in.retail.trade,
    Energy.and.water.supply..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Education..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Construction..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Repair.and.installation.of.machinery.and.equipment..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Consumer.Price.Index,
    Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay.
  )
# Primary model
aliased_formula <- as.formula(paste(var, "~ ."))
aliased_lm <- lm(aliased_formula, data = data_selected)
check_model_assumptions(aliased_lm)

```
"Total.offences" - ok after multicollinearity
"Offences.against.life" - fail after multicollinearity 
"Murder" - ok, normality is on edge
"Criminal.offenses.against.sexual.self.determination.in.total" - both fail
"Robberies.in.apartments" - ok
"Total.theft" - both fail
```{r}
# Reduce multicollinearity
offences_formula <- as.formula(paste(var, "~ . "))

reduced_formula <- reduce_multicollinearity(data_selected, offences_formula)

# Create model
reduced_lm <- lm(reduced_formula, data = data_selected)
check_model_assumptions(reduced_lm)

# Stepwise regression
sink(tempfile())
step_lm <- stepAIC(reduced_lm, direction = "both", data = data_selected)
sink()
check_model_assumptions(step_lm)
anova(aliased_lm, reduced_lm, step_lm)
#summary(aliased_lm)$adj.r.squared
#summary(reduced_lm)$adj.r.squared
#summary(step_lm)$adj.r.squared
cross_validate_lm(formula(step_lm), data_selected, 10)
```
RESULTS FOR EACH VARIABLE:
"Total.offences" - aliased_lm is the best fit, but step_lm has all assumptions
"Offences.against.life" - aliased_lm is the best fit - multicovarriant
"Murder" - weak fit, not a useful model
"Criminal.offenses.against.sexual.self.determination.in.total" - aliased_lm is the best fit - multicovarriant
"Robberies.in.apartments" - step_lm is the best
"Total.theft" - aliased_lm is the best fit - multicovarriant

ADJUSTED R-SQUARED FOR EACH VARIABLE FOR EACH MODEL   
"Total.offences" 
- 0.7334932 *
- 0.5955256
- 0.602088

"Offences.against.life" 
- 0.4734024 *
- 0.3663477
- 0.3653641

"Murder" 
- 0.3303478 
- 0.2837881
- 0.2807857 *

"Criminal.offenses.against.sexual.self.determination.in.total" 
- 0.7658117 *
- 0.5690452
- 0.5721362

"Robberies.in.apartments" 
- 0.5351619
- 0.5312519
- 0.5370938 *

"Total.theft" 
- 0.8501214 *
- 0.6355742
- 0.6383406

```{r}
# Total.offences manual multicollinearity prevention
data_selected <- data_cleaned %>%
  dplyr::select(
    Total.offences,
    Unemployment.rate,
    Vacancies.unfilled,
    Births_Total_x,
    Marriages.per.1000.inhabitants_x,
    Germans_Total_Departures.to.foreign.countries,
    Foreigners_Total_Arrivals.from.foreign.countries,
    Foreigners_Total_Departures.to.foreign.countries,
    Turnover.in.retail.trade,
    Energy.and.water.supply..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Education..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Construction..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Repair.and.installation.of.machinery.and.equipment..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Consumer.Price.Index,
    Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay.
  )
# Create model
manual_lm <- lm(Total.offences ~ .  -Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay. -Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay. -Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay. -Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay. -Vacancies.unfilled -Construction..Index.coll..agreed.monthly.earnings.with.sp.pay. -Energy.and.water.supply..Index.coll..agreed.monthly.earnings.with.sp.pay. -Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay., data = data_selected)
#summary(manual_lm)
#vif(manual_lm)
check_model_assumptions(manual_lm)

step_formula <- Total.offences ~ .  -Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay. -Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay. -Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay. -Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay. -Vacancies.unfilled -Construction..Index.coll..agreed.monthly.earnings.with.sp.pay. -Energy.and.water.supply..Index.coll..agreed.monthly.earnings.with.sp.pay. -Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay. -Education..Index.coll..agreed.monthly.earnings.with.sp.pay. -Repair.and.installation.of.machinery.and.equipment..Index.coll..agreed.monthly.earnings.with.sp.pay. -Marriages.per.1000.inhabitants_x

step_lm <- lm(step_formula, data = data_selected)
check_model_assumptions(step_lm)

anova(aliased_lm, reduced_lm, step_lm)
#summary(aliased_lm)$adj.r.squared
#summary(reduced_lm)$adj.r.squared
#summary(step_lm)$adj.r.squared

cross_validate_lm(step_formula, data_selected, 10)

```
Manual selection and multicovariate reduction:

Total.offences:
0.7334932
0.6777438 *
0.6694773 *

- variables:
Unemployment rate 
Births total
Germans total departures
Foreigners total arrivals
Foreigners total departures
Turnover in retail trade
Consumer price index


```{r}
# Criminal.offenses.against.sexual.self.determination.in.total manual multicollinearity prevention
data_selected <- data_cleaned %>%
  dplyr::select(
    Criminal.offenses.against.sexual.self.determination.in.total,
    Unemployment.rate,
    Vacancies.unfilled,
    Births_Total_x,
    Marriages.per.1000.inhabitants_x,
    Germans_Total_Departures.to.foreign.countries,
    Foreigners_Total_Arrivals.from.foreign.countries,
    Foreigners_Total_Departures.to.foreign.countries,
    Turnover.in.retail.trade,
    Energy.and.water.supply..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Education..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Construction..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Repair.and.installation.of.machinery.and.equipment..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Consumer.Price.Index,
    Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay.,
  )

aliased_lm <- lm(Criminal.offenses.against.sexual.self.determination.in.total ~ ., data = data_selected)
#check_model_assumptions(manual_lm)
# Create model
manual_lm <- lm(Criminal.offenses.against.sexual.self.determination.in.total ~ . -Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay. -Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay. -Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay. -Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay. -Construction..Index.coll..agreed.monthly.earnings.with.sp.pay. -Vacancies.unfilled -Education..Index.coll..agreed.monthly.earnings.with.sp.pay. -Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay., data = data_selected)

#summary(manual_lm)
#vif(manual_lm)
check_model_assumptions(manual_lm)

step_lm <- lm(Criminal.offenses.against.sexual.self.determination.in.total ~ . -Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay. -Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay. -Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay. -Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay. -Construction..Index.coll..agreed.monthly.earnings.with.sp.pay. -Vacancies.unfilled -Education..Index.coll..agreed.monthly.earnings.with.sp.pay. -Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay. -Marriages.per.1000.inhabitants_x -Foreigners_Total_Arrivals.from.foreign.countries , data = data_selected)
summary(step_lm)
```

```{r}
# Total.theft manual multicollinearity prevention
data_selected <- data_cleaned %>%
  dplyr::select(
    Total.theft,
    Unemployment.rate,
    Vacancies.unfilled,
    Births_Total_x,
    Marriages.per.1000.inhabitants_x,
    Germans_Total_Departures.to.foreign.countries,
    Foreigners_Total_Arrivals.from.foreign.countries,
    Foreigners_Total_Departures.to.foreign.countries,
    Turnover.in.retail.trade,
    Energy.and.water.supply..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Education..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Construction..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Repair.and.installation.of.machinery.and.equipment..Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay.,
    Consumer.Price.Index,
    Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay.,
  )

aliased_lm <- lm(Total.theft ~ ., data = data_selected)
#check_model_assumptions(manual_lm)
# Create model
manual_lm <- lm(Total.theft ~ . -Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay. -Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay. -Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay. -Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay. -Construction..Index.coll..agreed.monthly.earnings.with.sp.pay. -Education..Index.coll..agreed.monthly.earnings.with.sp.pay. -Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay. -Vacancies.unfilled -Germans_Total_Departures.to.foreign.countries, data = data_selected)

#vif(manual_lm)
check_model_assumptions(manual_lm)

step_lm <- lm(Total.theft ~ . -Overall.economy..Index.coll..agreed.monthly.earnings.with.sp.pay. -Transportation.and.storage..Index.coll..agreed.monthly.earnings.with.sp.pay. -Manufacturing..Index.coll..agreed.monthly.earnings.with.sp.pay. -Information.and.communication..Index.coll..agreed.monthly.earnings.with.sp.pay. -Construction..Index.coll..agreed.monthly.earnings.with.sp.pay. -Education..Index.coll..agreed.monthly.earnings.with.sp.pay. -Accommodation.and.food.and.beverage.service.act...Index.coll..agreed.monthly.earnings.with.sp.pay. -Vacancies.unfilled -Germans_Total_Departures.to.foreign.countries -Repair.and.installation.of.machinery.and.equipment..Index.coll..agreed.monthly.earnings.with.sp.pay. -Consumer.Price.Index, data = data_selected)

check_model_assumptions(step_lm)
```
VARIABLES PRESENT IN REDUCED MODELS

Economical variables:
Energy and water earnings ****
Information and communication earnings 
Education earnings 
Accomodation and food earnings 
Producers of capital goods 
Repair and installation of machinery ****
Turnover in Retail! ******
Unemployment rate ****
Vacancies count 
Consumer.Price.Index *****
Overall.economy

Social variables:
Immigration arrivals ***
Immigration departures **
German departures 
Births ****
Marriages! ******