---
title: "VISUALISATIONS"
output: html_document
date: "2024-12-29"
---

```{r setup, include=FALSE}
library(AER)
library(dplyr)       # all_of()
library(ggplot2)     # plotting  
library(tidyr)       # plotting
```
Ideas for observations - is there relationship between non-germans and specific crime types

basic percentage table - for selected crimes, are there any crimes which occur more often in non-german people
do these percentages coincide with crime spikes or are there othere associations (2016-2022 is declining, 2023 crime spiked)

YEARLY DATA ANALYSIS_______________________________________________________
```{r}
# Load data
data <- read.csv("final.csv")

#selection simple table
total_cases_table <- data %>%
  dplyr::select(
    total.offences..recorded.cases,
    bodily.injury..sects..223.227..229..231.PC...recorded.cases,
    offences.against.life..recorded.cases,
    murder..sect..211.PC...recorded.cases,
    robbery..extortion.resembling.robbery..and.assault.on.motorists.with.intent.to.rob..sects..249.252..255..316a.PC...recorded.cases,
    handbag.robbery..recorded.cases,
    bodily.injury..sects..223.227..229..231.PC...recorded.cases,
    serious.shoplifting..recorded.cases,
    street.crime..recorded.cases,
    unauthorised.trafficking..production.of..dispensing..and.possession.of.drugs..Sect..29a..1..no..2.NCA...significant.amount...recorded.cases,
    property.and.forgery.offences..recorded.cases
  )

# Second table: Select columns related to non-German suspects
non_german_suspects_table <- data %>%
  dplyr::select(
    total.offences..number.of.non.German.suspects,
    bodily.injury..sects..223.227..229..231.PC...number.of.non.German.suspects,
    offences.against.life..number.of.non.German.suspects,
    murder..sect..211.PC...number.of.non.German.suspects,
    robbery..extortion.resembling.robbery..and.assault.on.motorists.with.intent.to.rob..sects..249.252..255..316a.PC...number.of.non.German.suspects,
    handbag.robbery..number.of.non.German.suspects,
    bodily.injury..sects..223.227..229..231.PC...number.of.non.German.suspects,
    serious.shoplifting..number.of.non.German.suspects,
    street.crime..number.of.non.German.suspects,
    unauthorised.trafficking..production.of..dispensing..and.possession.of.drugs..Sect..29a..1..no..2.NCA...significant.amount...number.of.non.German.suspects,
    property.and.forgery.offences..number.of.non.German.suspects
  )

#what percentage of the crime was done by non_germans
non_german_percentage <- (non_german_suspects_table / total_cases_table) * 100
non_german_percentage$Year <- 2014:2023

```


VISUALISING FUNCTIONS__________________________
```{r}
visualise_stacked_bar <- function(dataset) {
    
  visualised_percentage_table <- dataset
  
  colnames(visualised_percentage_table) <- sapply(colnames(visualised_percentage_table), function(name) {
    if (grepl("\\.{2}", name)) {
      sub("\\.{2}.*", "", name)
    } else {
      name
    }
  })
  
  percentage_long <- visualised_percentage_table %>%
    pivot_longer(
      cols = -Year,  
      names_to = "Crime_Type", 
      values_to = "Percentage"  
    )
  
  ggplot(percentage_long, aes(x = Year, y = Percentage, fill = Crime_Type)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(
      title = "Distribution of Crimes Over the Years (2014-2023)",
      x = "Year",
      y = "Percentage of Total Crimes",
      fill = "Crime Type"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

visualise_crime_percentages_heatmap <- function(dataset) {
  
  colnames(dataset) <- sapply(colnames(dataset), function(name) {
    if (grepl("\\.{2}", name)) {
      sub("\\.{2}.*", "", name)
    } else {
      name
    }
  })
  
  percentage_long <- dataset %>%
    pivot_longer(
      cols = -Year,  
      names_to = "Crime_Type",  
      values_to = "Percentage"  
    )
  
  ggplot(percentage_long, aes(x = Year, y = Crime_Type, fill = Percentage)) +
    geom_tile(color = "white") +  
    geom_text(aes(label = round(Percentage, 1)), size = 3) +  
    scale_fill_gradient(low = "white", high = "red") +  
    labs(
      title = "Crime Percentages Heatmap (2014-2023)",
      x = "Year",
      y = "Crime Type",
      fill = "Percentage"
    ) +
    theme_minimal() +
    scale_x_continuous(breaks = 2014:2023) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),  
      axis.text.y = element_text(size = 10),  
      legend.position = "right"
    )
}
```


VISUALISING CRIME DISTRIBUTION BY MONTH__________________
```{r}
data_monthly <- read.csv("monthly.csv")
data_cleaned <- na.omit(data_monthly)
data_crime_months <- data_cleaned %>%
  dplyr::select(
    Date,
    Total.offences,
    Foreigners_Total_Arrivals.from.foreign.countries,
  )

data_crime_months$Date <- as.Date(data_crime_months$Date)

data_crime_months$Month <- format(data_crime_months$Date, "%b")  # Abbreviated month (Jan, Feb, etc.)
data_crime_months$Year <- format(data_crime_months$Date, "%Y")  # Extract year if needed


data_crime_long <- data_crime_months %>%
  pivot_longer(-c(Date, Month, Year), names_to = "Crime_Type", values_to = "Count")

ggplot(data_crime_long, aes(x = as.Date(Date), y = Count, color = Crime_Type)) +
  geom_line(size = 1) +
  labs(title = "Crime Counts Over Time", x = "Date", y = "Crime Count") +
  theme_minimal() +
  theme(legend.title = element_blank())

monthly_crime <- data_crime_long %>%
  group_by(Month, Crime_Type) %>%
  summarise(Average_Count = mean(Count))

czech_months <- c("led", "úno", "bře", "dub", "kvě", "čvn", "čvc", "srp", "zář", "říj", "lis", "pro")

monthly_crime <- monthly_crime %>%
  mutate(Month = factor(Month, levels = czech_months))

ggplot(monthly_crime, aes(x = Month, y = Average_Count, color = Crime_Type, group = Crime_Type)) +
  geom_line(size = 1) +
  geom_point(size = 2) +  
  labs(title = "Average Monthly Crime Fluctuations",
       x = "Month", y = "Average Crime Count") +
  theme_minimal() +
  theme(legend.title = element_blank())


```

VISUALISATION______________________________________
```{r}
visualise_crime_percentages_heatmap(non_german_percentage)

```
